---
html_document: default
author: "Savannah Cassis"
date: "January 29, 2020"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
pdf_document: default
title: "HW Extra Credit 1"
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Hypothesis: Does choice of super power depend on students political leaning?

#I want to test the hypothesis that whether students who choose flight or invisibility as their prefered super power varies with whether they are a Democrat, Republican, or Independent. This is a comparison of nominal to nominal scale data, so it'll require making a bar plot of frequencies:
  
  
  ## Analysis:

```{r}
rm(list = ls())
survey <-read.csv("http://www.courses.washington.edu/psy315/datasets/Psych315W20survey.csv")

# Then create the table 
myTable <- table(survey$vote,survey$superpower)

# The result is a table with both rows and columns, with labels:
myTable

# The labels can be pulled out using 'row.names' and 'colnames' (note 
# the inconsistency using '.' in the function names)
row.names(myTable)
colnames(myTable)

# The first and third rows correspond to Apple and PC's, and the 1st and 2nd columns
# correspond to left and right handedness.  This pulls out the relevant subset of rows and
# columns:

myTable <- myTable[c(1,3,5),c(1,2)]

```
## Results:

```{r}

# Here's the table of the results:
myTable

# And the bar graph:
barplot(myTable, 
        beside=TRUE,
        legend = row.names(myTable),
        col = c("Blue","Purple", "Red"))
```

## Summary Looking at the graph, it may seem that Republicans are slightly more likely to select invisibilty than flight, while Democrats are more likely to select flight. However, due to the majority of students voting democrat, over republican and independent combined, it is difficult to make a case because the sample is not sufficiently representative. 
